waRne - Putting cricket analytics in a spin

### 2020-10-03 paper_rob_working_good.Rmd
push
push

```{r libraries, include=FALSE}
library(tidyverse)
library(naniar)
library(visdat)
library(patchwork)

@@

```{r, echo=FALSE, warning=FALSE}
library(tidyverse)
cricjam_data_big %>% 
	dplyr::filter(batsman_id == 267192, 
								!is.na(bowling_style)) %>% 
	group_by(bowling_style, earned_runs) %>% 
	summarise(total = n()) %>% 
	group_by(bowling_style) %>% 
	mutate(totals = sum(total),
				 percent = total/totals, 
				 earned_runs = as.factor(earned_runs)) %>% 
	ggplot(aes(y = bowling_style, x = percent, color = earned_runs))+
	geom_point() +
	labs(color = "Runs", 
			 x = "% Balls Faced", 
			 y = "Bowling Style", 
			 title = "Steve Smith's Distribution of Runs Against Spin V Pace")

@@

```{r, echo=FALSE, warning=FALSE}
library(tidyverse)

cricjam_data_big %>% 
	dplyr::filter(batsman_id %in% rpo_players, 
								!is.na(pace_spin)) %>% 
	group_by(pace_spin, batsman_name) %>% 
	mutate(rpo = sum(runs)/n()*6) %>% 
	dplyr::summarise(mean_rpo = mean(rpo)) %>% 
	ggplot(aes(x = mean_rpo, y = batsman_name, color = pace_spin))+
	geom_point() +
	theme_minimal()+
	labs(x = "RPO", 
			 title = "RPO By Bowling Style", 
			 y = "", 
			 color = "")

```{r, echo=FALSE, warning=FALSE}

cricjam_data_big %>% 
	dplyr::filter(tournament_name %in% c("Twenty20 Big Bash", "Big Bash League"), 
								inning == 2) %>% 
	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
	group_by(match_id) %>% 
	slice(1) %>% 
	dplyr::select(batting_team_result, year) %>% 
	mutate(binary_result = if_else(batting_team_result=="win", 1, 0)) %>% 
	group_by(batting_team_result, year) %>% 
	summarise(win_count = n()) %>% 
	ungroup() %>% 
	pivot_wider(names_from = batting_team_result, values_from = win_count) %>% 
	rename(superover = 'super-over') %>% 
	replace_na(list(superover=0)) %>% 
	mutate(total_games = lose+win+superover, 
				 win_percentage = win/total_games, 
				 year = as.factor(year)) %>% 
	ggplot(aes(x = year, y = win_percentage, group = 1)) +
	geom_line(color = 'red')
```

@@

-   Do teams that bat second have a higher winning percentage than those who bat first?

@@

```{r, echo=FALSE, warning=FALSE}

n<-900
p<-0.5
qbinom(c(0.025,0.975),size=n,prob=p)

```{r, echo=FALSE, warning=FALSE}
df<-data.frame(
             Year = c(2010L,2011L,2012L,2013L,2014L,
                      2015L,2016L,2017L,2018L,2019L,2020L,2021L,2022L),
  number.of.heads = c(52L,55L,51L,70L,30L,40L,35L,
                      43L,49L,37L,39L,45L,32L)
)

library(broom)
df %>%
  # create variables to pass to binom.test
  mutate(n = 100, 
         x = n - number.of.heads, 
         p = 0.5,
         alternative = "two.sided",
         conf.level = 0.95) %>%
  # Run a binom.test for each row/year
  mutate(result = pmap(list(x, n,p), binom.test)) %>% 
  # Extract results from test into dataframe using broom::tidy
  mutate(result = map(result, tidy)) %>%
  mutate(result = map(result, ~select(.x, p.value, conf.low, conf.high))) %>% 
  unnest(cols = c(result)) %>%
  # plot the annual confidence intervals
  ggplot(aes(factor(Year), p.value)) + 
    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, color = (p.value > 0.05)), width=.1) +
  theme_minimal() +
  labs(y = "P(coin is fair)", x = "Year") +
  ggtitle("Annual coin fairness test")

@@

The yorker is considered the most difficult delivery to hit for a number of reason [Rob can you fill this in]. By pulling the delivery type from the play description on espncricinfo, every time a yorker was bowled we have the players who bowled and batted. We can take a look at the ability to hit yorkers for the closers we compared Dhoni to, and see how they all fare against the most difficult delivery in the game.

```{r}
middle_batsman_of_interest <- c(
  230559, #pollard
  32498, #yusef pathan
  321777, #david miller
  44936, #AB De Villiers
  276298, #andre russell
  8180, #shane watson
  34102, #rohit sharma
  28081, #Dhoni
  325026 #maxwell
  )

@@

## Easier engagement of fans

something something look towards how EPA has changed the way fans are engaged in NFL analytics - so maybe the change in run rate can be like EPA?

<https://twitter.com/cricvizanalyst/status/1311260989267087361>

@@

```{r, echo=FALSE}
library(tidyverse)

cricjam_data_big %>% 
	dplyr::filter(batsman_id %in% rpo_players, 
								!is.na(pace_spin)) %>% 
	group_by(pace_spin, batsman_name) %>% 
	mutate(rpo = sum(runs)/n()*6) %>% 
	dplyr::summarise(mean_rpo = mean(rpo)) %>% 
	ggplot(aes(x = mean_rpo, y = batsman_name, color = pace_spin))+
	geom_point() +
	theme_minimal()+
	labs(x = "RPO", 
			 title = "RPO By Bowling Style", 
			 y = "", 
			 color = "")

>>>>

```{r libraries, include=FALSE}
library(tidyverse)
library(naniar)
library(visdat)
library(patchwork)
library(cowplot)

@@

```{r, echo=FALSE, message=FALSE,warning=FALSE,out.width = "80%"}

p1<-cricjam_data_big %>% 
	dplyr::filter(batsman_id == 267192, 
								!is.na(bowling_style)) %>% 
	group_by(bowling_style, earned_runs) %>% 
	summarise(total = n()) %>% 
	group_by(bowling_style) %>% 
	mutate(totals = sum(total),
				 percent = total/totals, 
				 earned_runs = as.factor(earned_runs)) %>% 
	ggplot(aes(y = bowling_style, x = percent, color = earned_runs))+
	geom_point() +
	labs(color = "Runs", 
			 x = "% Balls Faced", 
			 y = "Bowling Style", 
			 title = "Steve Smith's Distribution of Runs Against Spin V Pace")

p2<-cricjam_data_big %>% 
	dplyr::filter(batsman_id %in% rpo_players, 
								!is.na(pace_spin)) %>% 
	group_by(pace_spin, batsman_name) %>% 
	mutate(rpo = sum(runs)/n()*6) %>% 
	dplyr::summarise(mean_rpo = mean(rpo)) %>% 
	ggplot(aes(x = mean_rpo, y = batsman_name, color = pace_spin))+
	geom_point() +
	theme_minimal()+
	labs(x = "RPO", 
			 title = "RPO By Bowling Style", 
			 y = "", 
			 color = "")

plot_grid(p1,p2, labels=c("",""))

@@

@@

-   Do teams that bat second have a different winning percentage than those who bat first?

@@

```{r, echo=FALSE, warning=FALSE}
library(broom)
cricjam_data_big %>% 
	dplyr::filter(inning == 1, 
								batting_team_result != "super-over") %>% 
	dplyr::filter(tournament_type == "Twenty20", 
                tournament_name %in% c("Twenty20 Big Bash", "Indian Premier League", "Big Bash League", 
                                       "Champions League Twenty20", "Pepsi Indian Premier League")) %>% 
	mutate(t_name = if_else(tournament_name %in% c("Big Bash League", "Twenty20 Big Bash"), "Big Bash", "IPL")) %>% 
	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
	group_by(year, batting_team_result, t_name) %>% 
	dplyr::select(batting_team_result, match_id, year, t_name) %>% 
	distinct() %>% 
	summarise(total = n()) %>% 
	pivot_wider(names_from = batting_team_result, values_from = total) %>% 
	ungroup() %>% 
	mutate(n =win+lose, 
				 p = 0.5, 
				 alternative = "two.sided", 
				 conf.level = 0.95, 
				 x = win) %>% 
	mutate(result = pmap(list(x, n,p), binom.test)) %>% 
  # Extract results from test into dataframe using broom::tidy
  mutate(result = map(result, tidy)) %>%
  mutate(result = map(result, ~select(.x, p.value, conf.low, conf.high))) %>% 
  unnest(cols = c(result)) %>%
  # plot the annual confidence intervals
  ggplot(aes(factor(year), p.value)) + 
    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, color = (p.value > 0.05)), width=.1) +
  theme_minimal() +
  labs(y = "P(Batting is fair)", x = "Year", color = "p > 0.05") +
  ggtitle("Annual 1st Inning Batting Fairness")+ 
  opts(axis.text.x=theme_text(angle=-90))+
	facet_wrap(~t_name)

@@

The yorker is considered the most difficult delivery to hit for a number of reasons . By pulling the delivery type from the play description on espncricinfo, every time a yorker was bowled we have the players who bowled and batted. We can take a look at the ability to hit yorkers for the closers we compared Dhoni to, and see how they all fare against the most difficult delivery in the game.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
middle_batsman_of_interest <- c(
  230559, #pollard
  32498, #yusef pathan
  321777, #david miller
  44936, #AB De Villiers
  276298, #andre russell
  8180, #shane watson
  34102, #rohit sharma
  28081, #Dhoni
  325026 #maxwell
  )

@@

@@

## Conclusion - Easier fan engagement

\newpage
